Control of Prosumer Networks

### 2015-10-28 Article.tex
update

%&latex
%%----------------------------------------------------------------------
%% ieeepes_skel.tex
%%
%% Skeleton file for papers for the IEEE Power Engineering Society using
%% package ieeepes.
%%
%% Not copyrighted. Copy this file to a different name and fill in your
%% text.
%%
%% Volker Kuhlmann
%% c/o EEE Dept
%% University of Canterbury
%% Private Bag 4800
%% Christchurch, New Zealand
%% Email: KUHLMAV@ELEC.CANTERBURY.AC.NZ
%%
% 1.3  13Apr99  Updated for ieeepes 4.0.
% 1.2  16Nov95  Fixed discussion, closure. Added summary.
% 1.1  12Nov95  Finished first release.
% 1.02 09Nov95  Option PStimes.
% 1.0  07Nov95  Created.
%%----------------------------------------------------------------------

\documentclass[10pt,twoside%
                ,draft%         % comment this out for final version
        ]{article}

\usepackage[%
        %       psphotos,%      % uncomment those options you want
        %       photofit,%
        %       draft,%
        %       PStimes,%
        ]{ieeepes}
\usepackage{algpseudocode}
\usepackage{algorithm}

\title{Control of Prosumer Networks}

\author{
        N. Gensollen\\
        Institut Mines Telecom\\
        Saclay, France
\and
        V. Gauthier\\
        Institut Mines Telecom\\
        Saclay, France
}

\begin{document}

% This \maketitle command is required from ieeepes version 4.0, to make
% ieeepes work correctly with newer LaTeX versions.
\maketitle

\begin{abstract}
The ability to control a given network by  dynamically injecting inputs offers ways to make sure that it stays in stable regions. In the case of smart grids, where end users exhibit complex behaviors and renewable production is quite unstable, control of the system  is paramount. In this paper, we study the case of a network composed of entities called prosumers. These agents have the ability to both consume and produce according to external conditions. We first model and simplify the underlying dynamic of such network using a second order coupled oscillators network model. Under some conditions, the system synchronize to a common frequency. However, in case of perturbations in the power distribution, the system might lose synchrony and require control to bring it back to the stable state. In this situation, control can be seen as energy  absorbed or injected in the system at specific locations. Moreover, the power outputs of the prosumers are susceptible to change such that loads and generators are not fixed. In this context, it is important to select the right subset
of nodes that yields, on average, the cheapest control in terms of energy.

\end{abstract}

\section{Introduction}

Modernizing the power grids and increasing the share of renewables in the production are two substantial objectives of the energetic transition. Because of parallel progresses in communication, data management, and storage, the upcoming emergence of a power grid "2.0", often called smart grid, appears as a cross-disciplinary challenge of the 21st century.

@@

These new agents that can both consume and produce are often called prosumers. This decentralized production is believed to lead to more resilient systems than today centralized power grid. Nonetheless, the difficulty of dealing with decentralized production is coupled with the high uncertainty of renewables. Ensure system stability in these conditions appears thus as a complex task.

@@

 In this paper, we model the prosumer network as a coupled oscillator network. Inspired by [], we use a second order Kuramoto model that can be seen as a simplified version of the underlying power grid dynamics (see Section II). We then use control theory in order to model the impact of the storage equipments on the network (see Section III). Using the relationships between submodular set functions
and the Gramian matrix [], we use a modified version of a greedy algorithm in order to find the smallest locations set that minimizes the average control energy while satisfying all the constraints introduced above (see section IV).
Finally, we show some results in Section V.\section{The Model}

In this section, we introduce the coupled oscillators network model used to simplify the power grid dynamic. The first part is largely inspired by [], and reproduced here for convenience of the reader.

@@

These non linear equations are obviously not very practical. We now use some approximations based on the fact that we consider only small deviations from the main frequency : $ \dot{ \delta}_i \sim \Omega $ which means that $ \omega_i = \dot{\theta_i} << \Omega $, such that the squared term $ K_D \dot{\theta}_i^2 $ can be neglected.
Moreover, we assume that the rate at which energy is stored in the kinetic term is much less of the rate at which energy is dissipated by friction, a common condition for mechanical systems : $ \ddot{ \theta }_i  << \dfrac{2 K_D}{I} $

@@

Where $ \alpha = \dfrac{2 K_D}{I} $ is the dissipation term, $ K_{ij} = \frac{P_{ij}^{MAX}}{I \Omega} $ are the coupling strengths, $ \psi_i = \left[ \frac{P_{S,i}}{I \Omega} - \frac{K_D \Omega}{I} \right] $, and $ g_{ij} $ is the coefficient of the adjacency matrix G.
By working in a rotating frame, we can simplify $ \psi_i=\frac{P_{S,i}}{I \Omega} $.

@@

\begin{equation}
\ddot{\theta} = \Psi - \alpha \dot{\theta} - KL\theta }
\end{equation}

@@

\begin{equation}
 Y(t+\Delta t) = A Y(t) }
\end{equation}

@@

\section{Control and Submodularity}

@@

In this section we provide a very brief introduction to what submodular functions are, and how their maximization could be achieved in reasonable time.

@@

Finding the set $ S_k $ of size k that maximizes a set function F is a difficult problem because the number of sets grows exponentially with the number of nodes. Therefore complete enumeration and evaluation is only a feasible solution on very small examples. Nevertheless, if the set function is submodular, a simple greedy heuristic returns a solution $ S_k^{\star} $ such that, in the worst case, $ \frac{F(S_k^{\star})}{F(S_k^{OPT})} \sim 63\%$ (where $ S_k^{OPT}$ is the optimum set of size k).

Since the evaluation of F can be costly, a well-known lazy-greedy variation has been proposed by Minoux in 19XX. This smart implementation uses the submodular structure of the marginal gains in order to reduce drastically the number of calls to F. This can result in speedups of several order of magnitudes.

@@

In this section, we introduce basic tools from control theory and relate them to the submodular functions introduced above. Recall that our objective is to maintain a given state in a dynamical system even if perturbations of the power distribution occur.

More precisely, we assume the control of a subset of the nodes that we call driver nodes. In our particular power grid situation, a driver node will be used to model a battery. This control assumption modifies the dynamic in the following way :

@@

Where the control matrix B is an indicator of the driver nodes, and $ u(t) $ represents the control vector input at time t. The system is said to be controllable in T steps is it can be steered from any initial state $ Y_0 $ to any final state $ Y_f $ through a sequence of control inputs $ u(t) $. If the system is controllable, there might be more than one sequence of control inputs that could realize such a thing. Among all these possibilities, we are interested in the one that requires the minimum amount of energy. Let $ \mathcal{E} = \int_{t_0}^{t_0+T} \Vert u(t) \Vert^2 dt $ be the energy used for the control of the system.

It can be shown that the control input sequence that minimizes the control energy can be written as :

\begin{equation}
 u^{\star}(t) = B^T(A^T)^{T-t-1}W^{-1}(T)\nu(T)
\end{equation}
where $ \nu(T) = Y_f - A^T Y_0 $ is the difference between the desired final state and the final free state, and $W(T) = \sum_{k=0}^{T} A^k B B^T (A^T)^k $ is called the Gramian matrix of the system. The Gramian is a popular tool in control theory since it is deeply linked to the controllability of the system. It can indeed be shown that the system is controllable if and only if the Gramian is not singular, and that its rank indicates the dimension of the controllable subspace. Besides, the minimum control energy associated with the inputs $ u^{\star}(t)$ can be written as :

@@

\section{Control of the Prosumer Network}

@@

\begin{algorithm}
\label{algo_1}
        \begin{algorithmic}
                \While {\small $ Not\ \left\{ \begin{array}{lll} rank \left[ W_{S_k}(T) \right] < 2N\ and \\ \forall t,\forall (i,j) \in V^2, g_{ij} \left| \theta_j(t) - \theta_i(t) \right| \leq 1\  and \\ \forall t,-\frac{\Lambda(0)}{I \Omega} \leq -\sum_{k=0}^{t}u^{\star}(k) \leq \frac{ \Lambda_{MAX} - \Lambda(0)}{I \Omega}and \\ \forall t,\ \left| u^{\star}(t) \right| \leq \frac{r}{I\Omega}\end{array} \right. $ }
    \State $k\ \ \ \ \ \ \ \ \ \ \ \gets k+1$
    \State $ S_k \ \ \ \ \ \ \ \ \ \gets Subopt(F, k)$
    \State $ u^{\star}(t)\ \ \ \ \ \ \gets OptControl(S_{k})$

        \end{algorithmic}
\end{algorithm}
\normalsize

\section{Results}
In this section, we provide some results in order to illustrate the work described above. We use a small 10 prosumers example, connected with a random topology. The line maximum capacities are selected  according to a normal distribution $\mathcal{N}(\mu_l,\sigma_l)$.  The power distribution is chosen as a zero mean normal distribution $ \mathcal{N}(0,\sigma_p) $, and the samples should satisfy the constraints \ref{synchro_constraint} and \ref{zero_sum_constraint}. Obviously, $\mu_l$, $\sigma_l$, and $\sigma_p$ should be chosen such that, without any perturbation, the lines are not already overloaded.

Once the synchronized state is reached, we apply a perturbation to the power distribution such that there is a mismatch between production and demand (constraint \ref{zero_sum_constraint} is not satisfied). We denote by $d$ the delay between the perturbation and the start of the control phase. $d$ depends on how fast the perturbation can be identified and located, the control inputs calculated and communicated without error through the communication network.

@@

Figure X illustrates the evolution of the state variables (phase angles and frequencies).
As visible, when d increases, the system state deviates more from the equilibrium and results in more energy being used to steer it back to synchrony. Figure XX shows how the global control energy evolves with the delay $ d$. The red dot indicates the maximum delay tolerable by the system before a constraint is broken. Here, some batteries become empty as shown in figure XXX.

We next investigate how the driver node set selected with algorithm X performs against other driver sets of the same size. Since, the example is relatively small, we can enumerate all these sets, and filter out the ones that do not meet all the constraints. We then sample the power distribution of the prosumers and compare the resulting mean  control energies for all the sets.

\bibliography{FilenameOfYourBibliography}

@@

%%
%% EOF ieeepes_skel.tex
%%----------------------------------------------------------------------

>>>>

\documentclass[conference]{IEEEtran}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Package
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{ifpdf}
\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.eps}
\fi

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[french, english]{babel}
\usepackage{cite}
\usepackage{caption}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{subcaption}
\usepackage{array}
\usepackage{float}
\usepackage{multicol}
\usepackage{algpseudocode}
\usepackage{algorithm}

\begin{document}
\pagestyle{plain}
\title{Control of Prosumer Networks}

\author{Nicolas Gensollen, Monique Becker, Vincent Gauthier and Michel Marot  \\
\IEEEauthorblockA{CNRS UMR 5157 SAMOVAR, \\
Telecom SudParis/Institut Mines Telecom\\
Email: \{nicolas.gensollen, vincent.gauthier, michel.marot, monique.becker\}@telecom-sudparis.eu}}

\maketitle

\begin{abstract}
The ability to control a given network by  dynamically injecting inputs offers ways to make sure that it stays in stable regions. In the case of smart grids, where end users exhibit complex behaviors and renewable production is quite unstable, control of the system  is paramount. In this paper, we study the case of a network composed of entities called prosumers. These agents have the ability to both consume and produce according to external conditions. We first model and simplify the underlying dynamic of such network using a second order coupled oscillators network model. Under some conditions, the system synchronize to a common frequency. However, in case of perturbations in the power distribution, the system might lose synchrony and require control to bring it back to the stable state. In this situation, control can be seen as energy  absorbed or injected in the system at specific locations. Moreover, the power outputs of the prosumers are susceptible to change such that loads and generators are not fixed. In this context, it is important to select the right subset of nodes that yields, on average, the cheapest control in terms of energy. 
\end{abstract}

\IEEEpeerreviewmaketitle

\section{Introduction}
\label{sec:introduction}

Modernizing the power grids and increasing the share of renewables in the production are two substantial objectives of the energetic transition. Because of parallel progresses in communication, data management, and storage, the upcoming emergence of a power grid "2.0", often called smart grid, appears as a cross-disciplinary challenge of the 21st century \cite{Ramchurn}.

@@

These new agents that can both consume and produce are often called prosumers \cite{Rathnayaka2012}. This decentralized production is believed to lead to more resilient systems than today centralized power grid. Nonetheless, the difficulty of dealing with decentralized production is coupled with the high uncertainty of renewables. Ensure system stability in these conditions appears thus as a complex task.

@@

 In this paper, we model the prosumer network as a coupled oscillator network. Inspired by \cite{Filatrella2008}, we use a second order Kuramoto model that can be seen as a simplified version of the underlying power grid dynamics (see Section \ref{sec:The_Model}). We then use control theory \cite{Liu2015} in order to model the impact of the storage equipments on the network (see Section \ref{sec:control_and_submodularity}). Using the relationships between submodular set functions and the Gramian matrix \cite{Summers2014}, we use a modified version of a greedy algorithm \cite{Minoux} in order to find the smallest locations set that minimizes the average control energy while satisfying all the constraints (see section \ref{sec:Control_prosumer}). Finally, we show some results in Section \ref{sec:Results}.

\section{The Model}
\label{sec:The_Model}

In this section, we introduce the coupled oscillators network model used to simplify the power grid dynamic. The detailed derivation of the second order dynamics can be found in \cite{Filatrella2008}, and the key points are reproduced here for the convenience of the reader.

@@

These non linear equations are obviously not very practical. We now use some approximations based on the fact that we consider only small deviations from the main frequency : $ \dot{ \delta}_i \sim \Omega $ which means that $ \omega_i = \dot{\theta_i} << \Omega $, such that the squared term $ K_D \dot{\theta}_i^2 $ can be neglected.
Moreover, we assume that the rate at which energy is stored in the kinetic term is much less of the rate at which energy is dissipated by friction, a common condition for mechanical systems : $ \ddot{ \theta }_i  << \frac{2 K_D}{I} $

@@

Where $ \alpha = \frac{2 K_D}{I} $ is the dissipation term, $ K_{ij} = \frac{P_{ij}^{MAX}}{I \Omega} $ are the coupling strengths, $ \psi_i = \left[ \frac{P_{S,i}}{I \Omega} - \frac{K_D \Omega}{I} \right] $, and $ g_{ij} $ is the coefficient of the adjacency matrix G.
By working in a rotating frame, we can simplify $ \psi_i=\frac{P_{S,i}}{I \Omega} $.

@@

\begin{equation}
\ddot{\theta} = \Psi - \alpha \dot{\theta} - KL\theta
\end{equation}

@@

\begin{equation}
 Y(t+\Delta t) = A Y(t)
\end{equation}

@@

\section{Control and Submodularity}
\label{sec:control_and_submodularity}

@@

In this section we provide a very brief introduction to what submodular functions are, and how their maximization could be achieved in reasonable time. More information can be found in \cite{Krause2014}.

@@

Finding the set $ S_k $ of size k that maximizes a set function F is a difficult problem because the number of sets grows exponentially with the number of nodes. Therefore complete enumeration and evaluation is only a feasible solution on very small examples. Nevertheless, if the set function is submodular, a simple greedy heuristic returns a solution $ S_k^{\star} $ such that, in the worst case, $ \frac{F(S_k^{\star})}{F(S_k^{OPT})} \sim 63\%$ (where $ S_k^{OPT}$ is the optimum set of size k) \cite{Krause2014}.

Since the evaluation of F can be costly, a well-known lazy-greedy variation has been proposed by \cite{Minoux}. This smart implementation uses the submodular structure of the marginal gains in order to reduce the number of calls to F. This can result in speedups of several order of magnitudes.

@@

In this section, we introduce basic tools from control theory and relate them to the submodular functions introduced above.

We assume the control of a subset of the nodes that we call driver nodes. In our particular power grid situation, a driver node can be seen as a prosumer where some additional equipment (battery, backup generator,...) will be used as to inject or absorb energy from the system. This control assumption modifies the system dynamic :

@@

Where the control matrix B is an indicator of the driver nodes, and $ u(t) $ represents the control vector input at time t. The system is said to be controllable in T steps if it can be steered from any initial state $ Y_0 $ to any final state $ Y_f $ through a sequence of control inputs $ u(t) $. If the system is controllable, there might be more than one sequence of control inputs that could achieve such a thing. Among all these possibilities, we are interested in the one that requires the minimum amount of energy \cite{Yan2012}. Let $ \mathcal{E} = \int_{t_0}^{t_0+T} \Vert u(t) \Vert^2 dt $ be the energy used for the control of the system.

It can be shown \cite{Liu2015} that the control input sequence that minimizes the control energy can be written as :

\begin{equation}
 u^{\star}(t) = B^T(A^T)^{T-t-1}W^{-1}(T)\nu(T)
\end{equation}

where $ \nu(T) = Y_f - A^T Y_0 $ is the difference between the desired final state and the final free state, and $W(T) = \sum_{k=0}^{T} A^k B B^T (A^T)^k $ is called the Gramian matrix of the system. The Gramian is a popular tool in control theory since it is deeply linked to the controllability of the system. It can indeed be shown that the system is controllable if and only if the Gramian is not singular, and that its rank indicates the dimension of the controllable subspace. Besides, the minimum control energy associated with the inputs $ u^{\star}(t)$ can be written as :

@@

\section{Control of the Prosumer Network}
\label{sec:Control_prosumer}

@@

\begin{algorithm}
	\label{algo_1}

        \begin{algorithmic}
                \While{  \small $ Not\ \left\{ \begin{array}{lll} 
                                              rank \left[ W_{S_k}(T) \right] < 2N\ and \\ 
                                              \forall t,\forall (i,j) \in V^2, g_{ij} \left| \theta_j(t) - \theta_i(t) \right| \leq 1\  and \\
                                              \forall t,-\frac{\Lambda(0)}{I \Omega} \leq -\sum_{k=0}^{t}u^{\star}(k) \leq \frac{ \Lambda_{MAX} - \Lambda(0)}{I \Omega}and \\ 
                                              \forall t,\ \left| u^{\star}(t) \right| \leq \frac{r}{I\Omega}

                                              \end{array} 
                                              \right. $ }
    \State $k\ \ \ \ \ \ \ \ \ \ \ \gets k+1$
    \State $ S_k \ \ \ \ \ \ \ \ \ \gets Subopt(F, k)$
    \State $ u^{\star}(t)\ \ \ \ \ \ \gets OptControl( S_{k} )$ 
\EndWhile
        \end{algorithmic}
\end{algorithm}
\normalsize

\section{Results}
\label{sec:Results}

In this section, we provide some results in order to illustrate the work described above. We use a small 10 prosumers example, connected with a random topology. The line maximum capacities are selected  according to a normal distribution $\mathcal{N}(\mu_l,\sigma_l)$.  The power distribution is chosen as a zero mean normal distribution $ \mathcal{N}(0,\sigma_p) $, and the samples should satisfy the constraints \ref{synchro_constraint} and \ref{zero_sum_constraint}. Obviously, $\mu_l$, $\sigma_l$, and $\sigma_p$ should be chosen such that, without any perturbation, the lines are not already overloaded.

\begin{figure}
\label{fig:frequecies}
\includegraphics[scale=.38]{frequencies}
\caption{Time series of the N frequencies $\omega_i$ (phase angles are not shown for clarity). $d$ is the delay between the perturbation and the begining of the control phase that bring the system bak in the synchronized state in T time steps. Here, $d$ was chosen quite large for the clarity of the figure. Note that, if the perturbation is still active at this point, the system still needs control in order to stay synchronized (batteries should still compensate the power imbalance)}
\end{figure}

Once the synchronized state is reached, we apply a perturbation to the power distribution such that there is a mismatch between production and demand (constraint \ref{zero_sum_constraint} is not satisfied). We denote by $d$ the delay between the perturbation and the start of the control phase (see figure \ref{fig:frequecies}). $d$ depends on how fast the perturbation can be identified and located, the control inputs calculated and communicated without error through the communication network.

@@

\begin{figure}
\label{fig:energy_vs_delay}
\includegraphics[scale=.38]{energy_vs_delay2}
\caption{ Control energy used in order to bring the system back to the synchronized state when the delay between perturbation and control increases, while the driver node set is fixed. The blue dotted curve shows the results for $ S_k^{\star} $ our algorithm \ref{algo_1}. The red line shows (average and standard deviation) the results when the driver set is sampled uniformly from all possible sets of the size $ k $.}
\end{figure}

As visible on figure \ref{fig:energy_vs_delay}, when d increases, the system deviates more from the equilibrium and this results in more energy being used to steer it back to synchrony. We also investigate here how the driver node set selected with algorithm \ref{algo_1} performs against other driver sets of the same size. we use a Monte-Carlo sampling of the drivers sets of the same size as our solution, and filter out the ones that do not meet all the constraints. Figure \ref{fig:energy_vs_delay} shows that our solution (blue dotted curve) uses less energy than other driver sets of the same size (red curve).

\section{Conclusion}
\label{sec:Conclusion}
C'est fini !

\bibliographystyle{IEEEtran}  
\bibliography{Article}

@@
